Forecasting U.S. Presidential Election Outcomes: A Poll-Based Predictive Model Using YouGov Data

### 2024-11-03 paper/paper.qmd
Final version.

The remainder of this paper is structured as follows: Section @sec-data provides an overview of the data, detailing its sources, cleaning procedures, and key variables like candidate name, political party, sample size, and polling end date. Section @sec-model discusses the model setup and justification, outlining the Bayesian linear regression approach and the rationale behind selecting specific predictors. Section @sec-result presents the results, examining the impact of party affiliation, candidate identity, and other factors on vote share predictions. Visualizations are included to illustrate key trends. Section @sec-disscussion explores broader implications, limitations, and potential extensions of this study, including sample size effects and directions for future research. Finally, sections @sec-appendix1 and @sec-appendix2 provide a detailed analysis of YouGov’s polling methodology and an idealized $100K survey methodology, offering practical insights for election forecasting.

@@

# Discussion {#sec-disscussion}
## Overview of Findings
In this paper, we built a predictive model to estimate the vote share of U.S. presidential candidates using polling data, with key predictors including political party, candidate name, sample size, and polling end date. The goal was to understand how these factors influence predicted vote shares, and to gain insights into the dynamics of political polling. The model’s high $R^2$ value (0.964) indicates it explains a substantial amount of variation in vote predictions, suggesting that it successfully captures relevant trends in the data. This outcome offers a deeper understanding of the factors driving poll-based vote share predictions in U.S. elections.

## Insights into Political Influence on Vote Share
One key finding from this analysis is the impact of political party on vote share predictions. Our model indicates that candidates from major parties, particularly Republicans, tend to receive higher predicted vote shares compared to third-party candidates. This is consistent with historical trends, where major-party candidates generally dominate U.S. elections. The model reinforces the idea that party affiliation is a powerful determinant in voter preferences, reflecting the established influence of major parties in American politics. This insight aligns with broader theories in political science, where party affiliation often serves as a shortcut for voters, helping them make quick decisions about candidates based on their general political orientation.

@@

## Weaknesses and Next Steps
While this model provides valuable insights, it also has some limitations. First, the model does not incorporate a variable for state-level differences, which could be significant in the U.S. electoral system, where the Electoral College, not the popular vote, ultimately determines the election outcome. A more detailed model could include state as a predictor or even focus on state-level vote shares to estimate Electoral College outcomes.

@@

In conclusion, while the current model provides a useful framework for analyzing vote share predictions, expanding its scope and incorporating new data sources could further enhance its predictive power and applicability. Understanding the nuances of voter behavior and poll-based predictions remains a critical area of research, especially in the context of modern elections where polling plays a central role in shaping public perception and media narratives.

@@

## Population, Frame, and Sample
The population targeted by YouGov’s polls comprises U.S. adult citizens, providing a frame that includes diverse demographics such as age, gender, race, income level, and political affiliation [@YouGov]. The samples are drawn from an online panel of respondents who voluntarily participate in YouGov’s surveys [@YouGov]. By leveraging an extensive online panel, YouGov strives to capture a broad, representative slice of public opinion on various political issues, including presidential election intent, candidate favorability, and policy priorities.

@@

## Questionnaire Design
YouGov’s questionnaires are carefully crafted to gather specific insights into voters’ opinions on candidates, political parties, and policy issues. The questions are standardized to allow comparison over time, aiding trend analysis. However, the online format of the questionnaire could impact responses, as participants are aware they are not in a monitored, controlled setting. Despite this limitation, YouGov’s design ensures clarity and consistency, although some complex topics may benefit from further simplification to enhance understanding across diverse education levels.

@@

## Summary
This survey methodology provides a comprehensive approach to election forecasting by leveraging stratified sampling, diverse recruitment, and thorough data validation. Structured with an engaging introductory section, well-ordered questions, and a respectful conclusion, this survey is designed to optimize response rates while yielding reliable insights into voter intentions and priorities. The iterative wave structure enables tracking of changes in sentiment over time, supporting a dynamic understanding of public opinion as the election date approaches.

>>>>

The remainder of this paper is structured as follows: Section @sec-data provides an overview of the data, detailing its sources, cleaning procedures, and key variables like candidate name, political party, sample size, and polling end date. Section @sec-model discusses the model setup and justification, outlining the Bayesian linear regression approach and the rationale behind selecting specific predictors. Section @sec-result presents the results, examining the impact of party affiliation, candidate identity, and other factors on vote share predictions. Visualizations are included to illustrate key trends. Section @sec-disscussion explores broader implications, limitations, and potential extensions of this study, including sample size effects and directions for future research. Finally, sections @sec-appendix1 and @sec-appendix2 provide a detailed analysis of YouGov’s polling methodology and an idealized $100K survey methodology, offering practical views for election forecasting.

@@

# Discussion {#sec-disscussion}
## Overview of Findings
In this paper, we built a predictive model to estimate the vote share of U.S. presidential candidates using polling data, with key predictors including political party, candidate name, sample size, and polling end date. The goal was to understand how these factors influence predicted vote shares, and to gain understanding on the dynamics of political polling. The model’s high $R^2$ value (0.964) indicates it explains a substantial amount of variation in vote predictions, suggesting that it successfully captures relevant trends in the data. This outcome offers a deeper understanding of the factors driving poll-based vote share predictions in U.S. elections.

## Political Influence on Vote Share
One key finding from this analysis is the impact of political party on vote share predictions. Our model indicates that candidates from major parties, particularly Republicans, tend to receive higher predicted vote shares compared to third-party candidates. This is consistent with historical trends, where major-party candidates generally dominate U.S. elections. The model reinforces the idea that party affiliation is a powerful determinant in voter preferences, reflecting the established influence of major parties in American politics. This finding aligns with broader theories in political science, where party affiliation often serves as a shortcut for voters, helping them make quick decisions about candidates based on their general political orientation.

@@

## Weaknesses and Next Steps
While this model provides valuable findings, it also has some limitations. First, the model does not incorporate a variable for state-level differences, which could be significant in the U.S. electoral system, where the Electoral College, not the popular vote, ultimately determines the election outcome. A more detailed model could include state as a predictor or even focus on state-level vote shares to estimate Electoral College outcomes.

@@

In conclusion, while the current model provides a useful framework for analyzing vote share predictions, expanding its scope and incorporating new data sources could further enhance its predictive power and applicability. Understanding the nuances of voter behavior and poll-based predictions remains an important area of research, especially in the context of modern elections where polling plays a central role in shaping public perception and media narratives.

@@

## Population, Frame, and Sample
The population targeted by YouGov’s polls comprises U.S. adult citizens, providing a frame that includes diverse demographics such as age, gender, race, income level, and political affiliation [@YouGov]. The samples are drawn from an online panel of respondents who voluntarily participate in YouGov’s surveys [@YouGov]. By using an extensive online panel, YouGov strives to capture a broad, representative slice of public opinion on various political issues, including presidential election intent, candidate favorability, and policy priorities.

@@

## Questionnaire Design
YouGov’s questionnaires are carefully crafted to gather specific information about voters’ opinions on candidates, political parties, and policy issues. The questions are standardized to allow comparison over time, aiding trend analysis. However, the online format of the questionnaire could impact responses, as participants are aware they are not in a monitored, controlled setting. Despite this limitation, YouGov’s design ensures clarity and consistency, although some complex topics may benefit from further simplification to enhance understanding across diverse education levels.

@@

## Summary
This survey methodology provides a comprehensive approach to election forecasting by using stratified sampling, diverse recruitment, and thorough data validation. Structured with an engaging introductory section, well-ordered questions, and a respectful conclusion, this survey is designed to optimize response rates while yielding reliable findings on voter intentions and priorities. The iterative wave structure enables tracking of changes in sentiment over time, supporting a dynamic understanding of public opinion as the election date approaches.
